Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2020-12-03 mice_synthesizing/paper.Rmd
some changes

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Because collecting research data requires an enormous investment both in terms of time and monetary resources, openly accessible research data bears the potential of increasing the scientific returns. Additionally, the fact that public funds are used for data collection results in increasing demand for the collected data. Nevertheless, the possibilities to share research data directly are often very limited due to restrictions with regard to privacy and confidentiality. Simply anonimizing the data is not enough to fulfil these requirements [@ohm_broken_2009]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, thereby drastically reducing the data quality and the usability for further research.

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the general idea is to use actually observed data to generate multiply imputed synthetic datasets that can be freely disclosed. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. Using this approach, the researcher could replace the observed data set as a whole with multiple synthetic versions. Alternatively, the researcher could opt to only replace a subset of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available data sets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, say, income values above a certain threshold.

@@

Potentially, this approach could fulfil the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data is highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these high-quality data sets should be kept at a minimum.

@@

The remainder of this paper is organized as follows. First, we will shortly describe the `mice` algorithm for the creation of synthetic data that reassures the privacy and confidentiality of the participants. Second, we will demonstrate a straightforward workflow for imputation of synthetic data with `mice`. The third step is to summarize the variance estimators that have been proposed in the past. Finally, we will assess which variance estimator provides valid inferences from the synthetic data.

@@

The `mice` package has been developed originally for multiple imputation of missing data. In this context, the aim is to replace missing values due to nonresponse by plausible values from the posterior predictive distribution of the variable containing the missings. Doing so, `mice` makes use of fully conditional specification (FCS), which breaks down the multivariate distribution of the data $\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})$ into $j = 1, 2, \dots, k$ univariate conditional densities, with $k$ the number of variables in the data. Using FCS, for every variable containing missing values a model is constructed, to impute the missing values $Y_{j, mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ on a variable by variable basis. Note that the predictor matrix $Y_{-j}$ may contain yet imputed values from an earlier imputation step, and thus will be updated after every iteration. This procedure is applied $m$ times, resulting in $m$ completed datasets $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$, with $\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})$. __MAYBE ADD AN EXAMPLE OF HOW TO DO IT IN `mice`__

__DISCUSS ESTIMATORS FOR MISSING DATA HERE, OR AT A LATER POINT__

>>>>

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. For example, collecting research data requires an enormous investment both in terms of time and monetary resources. Openly accessible research data bears the potential of increasing the scientific returns for the same data collection effort. Additionally, the fact that public funds are used for data collection results in increasing demand for the collected data. Nevertheless, the possibilities to share research data directly are often very limited due to restrictions with regard to privacy and confidentiality.

A quick approach to lessen the identities in the data is anonimization. However, simply anonimizing the data is not sufficient to fulfill contemporary privacy and confidentiality  requirements [@ohm_broken_2009]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, thereby reducing the data quality and the scientific return for re-using the same data for further research.

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the general idea is to use actually observed data to generate multiply imputed synthetic data sets that can be freely disclosed. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. Using this approach, the researcher could replace the observed data set as a whole with multiple synthetic versions. Alternatively, the researcher could opt to only replace a subset of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available data sets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, such as income values above a certain threshold. [**aren't synthetic values then not also too disclosive?**]

@@

Potentially, this approach could fulfill the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data is highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these high-quality data sets should be kept at a minimum.

@@

The remainder of this paper is organized as follows. First, we will shortly describe the `mice` algorithm for the creation of synthetic data. The aim is to generate synthetic sets that reassure the privacy and confidentiality of the participants. Second, we will demonstrate a straightforward workflow for imputation of synthetic data with `mice`. The third step is to summarize the variance estimators that have been proposed in the past. Finally, we will assess which variance estimator provides valid inferences from the synthetic data.

@@

The `mice` package has been developed for multiple imputation of missing data. In this context, the aim is to replace missing values due to nonresponse by plausible values from the posterior predictive distribution of the variable containing the missings. Doing so, `mice` makes use of fully conditional specification (FCS) [**ref to Stef's paper in 2006**], which breaks down the multivariate distribution of the data $\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})$ into $j = 1, 2, \dots, k$ univariate conditional densities, with $k$ the number of variables in the data. Using FCS, for every variable containing missing values a model is constructed, to impute the missing values $Y_{j, mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ on a variable by variable basis. Note that the predictor matrix $Y_{-j}$ may contain yet imputed values from an earlier imputation step, and thus will be updated after every iteration. This procedure is applied $m$ times, resulting in $m$ completed datasets $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$, with $\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})$. __MAYBE ADD AN EXAMPLE OF HOW TO DO IT IN `mice`__

__DISCUSS ESTIMATORS FOR MISSING AT A LATER POINT__
